### 2019-05-24 reviews/response.md

We thank the reviewers for their detailed and insightful feedback.

We have added four new implementations to Task Bench: Dask, bulk synchronous MPI,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section, including preliminary results from experiments with GPUs on
the Piz Daint supercomputer.

Major changes in the revised paper:

  * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
  * New sections:
      - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
      - Evaluation: Section 5.9 (GPU experiment)
  * New figures:
      - Listings 1 and 2 (code samples)
      - Figure 15 (GPU experiment)
  * Updated Figures 3 and 4 (see also lines 303-315)
  * Updated Section 7 with insights derived from Task Bench

@@

Reviewer 2:

Figures 3 and 4 have been updated and new text added.

Reviewer 3:

Weak/strong scaling and TPS are also application and system dependent
metrics. METG differs from these metrics in that can simultaneously
(a) capture the behavior of the system at the limit of its scalability
(which is not true of weak/strong scaling) and (b) while achieving a
specified amount of useful work (which is not true of TPS). METG also
isolates the impact of changes to the communication topology at scale,
whereas in a strong scaling study this effect would be entangled with
issues of system overhead. Examples of how to compute METG from
application wall time have been added to Section 2.

### 2019-05-24 4_implementation.tex
Hide weak/strong scaling figures and fiddle with figure placement.
More changes to shrink text.

Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
\emph{chares}, are objects in their own address space.
Chares communicate data and synchronize via messages and can be moved
to balance load. Our Task Bench implementation uses a chare
array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
dependencies are all available. Collectives synchronize the
completion of the task graph execution.

@@

OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
for OpenMP is very similar to OmpSs and uses tasks with
task dependencies. We tested both GNU GOMP and Intel KMP
implementations of OpenMP and found KMP to be better performing on Cori.

@@

Tasks in Realm are executed in a deferred manner, with dependencies
between tasks explicitly defined by \emph{events} passed from one task
to another. Realm's data model supports collections that live in a
specific memory. Data must be explicitly copied, and the copies
synchronized with tasks via events.

Realm also supports a \emph{subgraph} API that enables additional
optimizations. We report numbers for the subgraph version of Task
Bench as it provides better performance.

>>>>

Charm++~\cite{Charmpp93} is an explicitly parallel actor-based programming system. Actors, or
\emph{chares}, are objects in their own address space.
Chares communicate data and synchronize via messages and can be moved
to balance load. Our Task Bench implementation uses a chare
array for the task graph, with one chare for each column. Messages implement dependencies; a task executes as soon as its
dependencies are all available.

%% Collectives synchronize the
%% completion of the task graph execution.

@@

OpenMP~\cite{OpenMPSpec40} is the industry standard API for loop-based
parallelism on shared-memory systems, and supports task dependencies as of version 4.0. Our Task Bench implementation
for OpenMP is similar to OmpSs and uses
task dependencies. We tested with GNU GOMP and Intel KMP
and found KMP to be better performing.

@@

Tasks in Realm are asynchronous, with dependencies
defined by \emph{events} passed from one task
to another. Realm's data model supports collections that live in a
specific memory. Data must be explicitly copied, and the copies
synchronized with tasks via events.
Realm also supports a \emph{subgraph} API that enables additional
optimizations. We use subgraphs in the Task
Bench implementation.

### 2019-05-24 5_evaluation.tex
Hide weak/strong scaling figures and fiddle with figure placement.
More changes to shrink text.

% FIXME: get Piz Daint software versions.

@@

@@

% Really want this to appear in the next section, but Latex forces it to be placed here to get the layout we want.
\input{e3_metg_compute}

@@

Spark, Swift/T and TensorFlow
are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems would have required
unreasonably large problem sizes (beyond what can be completed in 6
hours) to reach 50\% efficiency.

@@

\input{e8_weak}
\input{e9_strong}
\input{e5_imbalance}

@@

\subsection{Comparison with Weak and Strong Scaling}

@@

Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
conventional strong and weak scaling of the stencil pattern for
comparison with the METG results in
Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
average task granularity is larger than the METG for most systems, the
performance is ideal. For several systems METG rises with node count,
and this becomes visible in the weak and strong scaling graphs as
non-ideal performance. Thus the shape of the weak and strong scaling
graphs effectively interpolates between ideal performance and the METG
curve depending on how close each system is to being limited by
overhead.

@@

The difference between the remaining systems is due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at very small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers here because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave the analysis of persisitent load imbalance to future work.

>>>>

@@

\input{e3_metg_compute}

@@

@@

Spark, Swift/T and TensorFlow
are omitted from comparisons with more complicated dependencies, as the higher overheads of these systems require
unreasonably large problem sizes (beyond what can be completed in 6
hours) to reach 50\% efficiency.

@@

\input{e5_imbalance}

@@

\zap{
\input{e8_weak}
\input{e9_strong}
\subsection{Comparison with Weak and Strong Scaling}

Figures~\ref{fig:weak-scaling} and \ref{fig:strong-scaling} show
conventional strong and weak scaling of the stencil pattern for
comparison with the METG results in
Figure~\ref{fig:metg-compute-stencil}. At small node counts, where the
average task granularity is larger than the METG for most systems, the
performance is ideal. For several systems METG rises with node count,
and this becomes visible in the weak and strong scaling graphs as
non-ideal performance. Thus the shape of the weak and strong scaling
graphs effectively interpolates between ideal performance and the METG
curve depending on how close each system is to being limited by
overhead.
} % zap

@@

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persisitent load imbalance to future work.
